Causes and effects in Dichotomous Comparative Judgments: an information-theoretical system of plausible mechanism

### 2024-10-18 _tex/index.tex
Built site for gh-pages

Over the past decade, numerous studies have documented the effectiveness
of the \emph{comparative judgment} (CJ) method \citep{Thurstone_1927}
for assessing competencies and traits. These studies have evaluated CJ
from two main perspectives: its ability to produce reliable and valid
trait scores, and its practical applicability. In terms of reliability
and validity, research has shown that CJ can generate precise and
consistent scores that accurately represent the traits being measured.
Notable contributions in this research area include studies by Pollit
\citetext{\citeyear{Pollitt_2012a}; \citeyear{Pollitt_2012b}},
\citet{Whitehouse_2012}, \citet{vanDaal_et_al_2016},
\citet{Lesterhuis_2018}, \citet{Bramley_et_al_2019},
\citet{Verhavert_et_al_2019}, \citet{Crompvoets_et_al_2022}, and
\citet{Bouwer_et_al_2023}. Regarding practical applicability, several
studies have highlighted the method's versatility in both educational
and non-educational contexts, presenting it as an efficient and
effective alternative for measurement and evaluation. Key examples in
this research area include the works of \citet{Jones_2015},
\citet{Bartholomew_et_al_2018}, \citet{Jones_et_al_2019},
\citet{Marshall_et_al_2020}, \citet{Bartholomew_et_al_2020}, and
\citet{Boonen_et_al_2020}.

Despite the growing number of CJ studies, the unsystematic and
fragmented research approaches employed in the literature have
overlooked several critical issues concerning CJ. These issues can be
categorized into concerns about the method's structural model,
measurement model, and experimental design.

A key issue in the first category is the apparent disconnect between the
method's structural and measurement models. A common practice in CJ
literature involves performing data analysis and hypothesis testing on
scores previously estimated using the Bradley-Terry-Luce (BTL) model
\citep{Bradley_et_al_1952, Luce_1959}. These studies use the scores or
their transformations to identify `misfit' judges and stimuli
\citep{Pollitt_2012b, vanDaal_et_al_2017}, detect `bias' in judges'
ratings \citep{Pollitt_et_al_2003, Pollitt_2012b}, or test various
hypothesis
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.
However, since CJ scores are parameter estimates with inherent
uncertainty, the statistical literature suggests that separating the
analysis from this uncertainty may artificially inflate the precision
and power of the results. This, in turn, could increase the risk of
committing a type I error, where a null hypothesis is wrongly rejected
\citep{McElreath_2020}.

>>>>

Over the past decade, numerous studies have documented the effectiveness
of the \emph{comparative judgment} (CJ) method \citep{Thurstone_1927}
for assessing competencies and traits. These studies have evaluated CJ
from two main perspectives: its ability to produce reliable and valid
trait scores, and its practical applicability. Research on reliability
and validity has shown that CJ can generate precise and consistent
scores that accurately represent the traits being measured
\citep{Pollitt_2012a, Pollitt_2012b, Whitehouse_2012, vanDaal_et_al_2016, Lesterhuis_2018, Bramley_et_al_2019, Verhavert_et_al_2019, Crompvoets_et_al_2022, Bouwer_et_al_2023}.
Regarding practical applicability, several studies have highlighted CJ's
versatility across both educational and non-educational contexts,
presenting it as an efficient and effective alternative for measurement
and evaluation
\citep{Jones_2015, Bartholomew_et_al_2018, Jones_et_al_2019, Marshall_et_al_2020, Bartholomew_et_al_2020, Boonen_et_al_2020}.

Despite the growing number of CJ studies, the unsystematic and
fragmented research approaches employed in the literature have
overlooked several critical issues concerning the method. These issues
fall into three main categories: concerns about the structural model,
the measurement model, and the experimental design of CJ. In the
following sections, each issue will be discussed in detail, followed by
the introduction of an approach that addresses all three concerns
simultaneously.

A key issue in the first category is the apparent disconnect between the
method's structural and measurement models. In CJ literature, it is
common to perform data analysis and hypothesis testing on scores
estimated using the Bradley-Terry-Luce (BTL) model
\citep{Bradley_et_al_1952, Luce_1959}. Several studies use the scores
generated by the BTL model or their transformations to identify `misfit'
judges and stimuli \citep{Pollitt_2012b, vanDaal_et_al_2017}, detect
`bias' in judges' ratings \citep{Pollitt_et_al_2003, Pollitt_2012b}, or
test various hypotheses about the underlying trait being measured
\citep{Bramley_et_al_2019, Boonen_et_al_2020, Bouwer_et_al_2023, vanDaal_et_al_2017, Jones_et_al_2019, Gijsen_et_al_2021}.
However, since these scores are parameter estimates with inherent
uncertainty, the statistical literature suggests that separating the
analysis from this uncertainty may artificially inflate the precision
and power of the results. This, in turn, could increase the risk of
committing a type I error, where a null hypothesis is wrongly rejected
\citep{McElreath_2020}. To address this issue properly, the approach
should follow a strategy similar to that used in Structural Equation
Modeling (SEM), where data analysis and hypothesis testing occur at the
structural model level, while the BTL model functions as the measurement
model.
